---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***The package is in an early development stage and was posted to facilitate collaboration.***

[![Build Status](https://travis-ci.org/konradedgar/SmarterScotland.svg?branch=master)](https://travis-ci.org/konradedgar/SmarterScotland) [![codecov](https://codecov.io/gh/konradedgar/SmarterScotland/branch/master/graph/badge.svg)](https://codecov.io/gh/konradedgar/SmarterScotland) [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/konradedgar/SmarterScotland?branch=master&svg=true)](https://ci.appveyor.com/project/konradedgar/SmarterScotland)


## Overview

The SmarterScotland package provides a convenient interface for sourcing and interrogating data available through the [statistics.gov.scot](http://statistics.gov.scot).

## Installation

```{r installation, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
if ("devtools" %in% rownames(installed.packages())) {
  devtools::install_github(repo = "konradedgar/SmarterScotland", build_vignettes = TRUE)
} else {
  install.packages("devtools")
  devtools::install_github(repo ="konradedgar/SmarterScotland", build_vignettes = TRUE)
}
```

## Other projects

For better and more mature, similar projects have a look at:

* [SmarterPoland](https://cran.r-project.org/web/packages/SmarterPoland/index.html)
* [eurostat](https://cran.r-project.org/web/packages/eurostat/index.html)