---
title: "Basic Usage"
author: "Konrad Zdeb"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Load package
require("SmarterScotland")
```

# Introduction

The following packages makes use of the SPARQL [^1] language queries in order to query the data available through the [statistics.gov.scot](http://statistics.gov.scot) portal [^2]. In effect, the app works as a wrapper for SPARQL queries providing a suite of common functionalities relevant to how the Scottish public data is being utilises.

[^1]: Detailed information on the SPARQL is available through [W3C pages](https://www.w3.org/TR/rdf-sparql-query/).
[^2]: The implementation follows part of the *Open Data Strategy* implemented by the Scottish Government, policy papers and relevant legislation is available through [Scottish Government pages](http://www.gov.scot/Topics/Economy/digital/digitalservices/datamanagement/OpenData).

## What data is available

The function `get_ToC` provides a simple mechanism generating information on the initial data that is available through the [statistics.gov.scot](http://statistics.gov.scot). For instance the query below will return a list of indicators for which the keyword *pupil* was found.

```{r get_ToC_example1}
head(get_ToC(which = "pupil"))
```

A more informative approach to sourcing the relevant data relies on producing a matrix of indicators available across specific geographies. This is shown below.

# Sample queries

## Available geographies

The list of available geographies can be sourced with use of the query below.
