---
title: "Basic Workflow"
author: "Konrad Zdeb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)

# Enable httptest for vignette development
library(httptest)
start_vignette("basic-workflow")

# Load package
library("SmarterScotland")
```

# Introduction

The following packages makes use of the SPARQL[^1] language queries in order to query the data available through the [statistics.gov.scot](http://statistics.gov.scot) portal[^2]. In effect, the app works as a wrapper for SPARQL queries providing a suite of common functionalities relevant to how the Scottish public data is being utilises.

# Workflow

The proposed workflow looks at obtaning *life expectancy* data for the *Glasgow City.* The workflow focuses on the following steps:

1. **Data Selection:** Identification of the relevant data
2. **Geography Selection:** Veryfying name of geogrphy for which we want to source data
3. **Data Sourcing:** Querying [statistics.gov.scot](http://statistics.gov.scot) SPARQL endpoint
4. **Pre-processing and Analysis:** Fixing data types and removing redundant columns

## 1. Data Selection

The package comes with an embedded data sets containing list of data and geographies available through the [statistics.gov.scot](http://statistics.gov.scot) endpoint. To see the list of available data sets refer to the `data("available_data_sets")` objects. The `available_data_sets` contains `subject` column that can be used to identify relevant data, as shown below.

```{r get_available_data, eval=FALSE}
data("available_data_sets")
subset.data.frame(x = available_data_sets, 
                  subset = grepl(pattern = "expectancy", x = subject))
```
```{r show_available_data, echo=FALSE}
knitr::kable(
  subset.data.frame(x = available_data_sets, 
                  subset = grepl(pattern = "expectancy", x = subject)),
  caption = "Available Life Expectancy Data"
)
```

## 2. Geography

The object `data("standard_geography_code_register")` is provided to facilitate veryfying existence of the geography we are interested in. Similarly, as in the previous step we can use `grep()` to verify existence of that object in the provided data frame. For that purpose we can check the `value` column for existence of the desired string.

```{r show_available_geographies, eval=FALSE}
data("standard_geography_code_register")
subset.data.frame(x = standard_geography_code_register,
                  subset = grepl(pattern = "^Glasgow City$", x = value))
```
```{r table_available_geographies, echo=FALSE}
knitr::kable(
  subset.data.frame(x = standard_geography_code_register,
                  subset = grepl(pattern = "^Glasgow City$", x = value))
)
```

## 3. Data Sourcing

Data sourceing is achieved through `get_geography_data()` function. Following from the example below the function can be applied as follows

```{r}
dta_Glasgow_Life_Expectancy <- get_geography_data(data_set = "life-expectancy",
                                                  geography = "Glasgow City",
                                                  measure = "count"
                                                  pre_process_results = TRUE)
```

## 4. Pre-processing and Analysis

The pre-processing of the results already took place and was achieved by passing `pre_process_results = TRUE` to `get_geography_data()` function. The results can be also pre-processed independently by calling `pre_process_data()` on a returned object.

# Conclusion

For the efficiency the objects `data("available_data_sets")` and `data("standard_geography_code_register")` are provided in the package. Those objects are not updated too freqntely and providing them as statics data sets makes sense from the efficiency perspective. It is however possible to source live the mentioned objects. This is discussed in `vignette("support-functions")`.

[^1]: Detailed information on the SPARQL is available through [W3C pages](https://www.w3.org/TR/rdf-sparql-query/).
[^2]: The implementation follows part of the *Open Data Strategy* implemented by the Scottish Government that can be accessed through the [Scottish Government pages](http://www.gov.scot/Topics/Economy/digital/digitalservices/datamanagement/OpenData).
[^3]: This is currently being developed; this behaviour will be controlled by `SmarterScotland.suffix_rdf` option.

```{r end-setup, include=FALSE}
end_vignette()
```