---
title: "Other Functions"
author: "Konrad Zdeb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Other Functions}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(httptest)
start_vignette("other-functions")

# Load package
library("SmarterScotland")
```

# Introduction 


# Exploring [statistics.gov.scot](http://statistics.gov.scot)

## Data sets

Additionally, the function `get_ToC` provides a simple mechanism generating information on the initial data that is available through the [statistics.gov.scot](http://statistics.gov.scot). For instance the query below will return a list of indicators for which the keyword *pupil* was found. Expectedly, the returned data sets will mostly correspond to educational attainment. The query sources available data via SPARQL end point and does not rely on stored data sets.

```{r get_available_data_example}
available_data_education <- get_available_datasets(pattern = "pupil")
knitr::kable(available_data_education, row.names = FALSE)
```

## Geographies



```{r end-setup, include=FALSE}
end_vignette()
```
