---
title: "Support Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Enable httptest for vignette development
library(httptest)
start_vignette("support-functions")

# Load package
library("SmarterScotland")
```

```{r setup}
library(SmarterScotland)
```

The `vignette("basic-workflow")` vignette provides an outline of basic functionalities 

# Live reference data

More recent versions of the provided data sets can be sourced live from the [statistics.gov.scot](http://statistics.gov.scot) portal using dedicated functions, `get_available_datasets` sources available data and `get_available_geographies()` sources and parses Standard Geography Code Register. Searching for thematic data sets can be achieved simply using queries, as shown below. 

```{r search_jsa_data, results='hide'}
get_available_datasets(pattern = "expectancy")
```
```{r search_jsa_data_results, echo=FALSE}
knitr::kable(get_available_datasets(pattern = "expectancy"))
```

Additional information on the available data can be gathered by following the links returned in the data set above. For instance [http://statistics.gov.scot/data/pupil-attainment-asn](http://statistics.gov.scot/data/pupil-attainment-asn) points to *Pupil Attainment (English and Maths)* data set.